Naturalness and Long-Lived Beyond the Standard Model Particles

### 2017-08-05 tex/draft.tex
Dumb down SM section

\newpage
\tableofcontents
\newpage
\doublespacing
\linenumbers
\pagenumbering{arabic}

% Things to check
\fxnote{When should symbols be italicized and when should they be mathrm'd?} 
\fxnote{fine-tuning or fine tuning?}

\section{The Standard Model} \label{SM}
    The standard model of particle physics (SM) describes all known particles and their non-gravitational interactions and is experimentally verified to astonishing \fxnote{weak word} accuracy. The SM consists of \num{12} spin-$\frac{1}{2}$ particles, the fermions, that make up all observed matter; \num{12} spin-1 particles, the gauge bosons, that communicate the electromagnetic, weak, and strong forces; and one fundamental scalar, the Higgs boson, which will be described in detail below.

    All SM particles transform according to the gauged symmetry group\fxnote{be good at describing the meaning of `gauged symmetry group'} $\SUthreecSUtwoLUoneY$, where $\SUthreec$ is associated with color charge, $\SUtwoL$ is associated with weak isospin, and $\UoneY$ is associated with weak hypercharge. The gauge bosons are associated with the generators of these groups, thus there are eight gluons associated with $\SUthreec$, three W bosons associated with $\SUtwoL$, and one B boson associated with $\UoneY$. The physical gauge bosons, which are linear combinations of the Ws and B, are the $\Wp$, $\Wm$, $\Z$, and photon ($\gamma$). The fermions are categorized by which representation of each group they fit into\fxnote{make sure I'm good at explaining this}. For example, $\SUthreec$ distinguishes the quarks (which live in the fundamental representation) from the leptons (which are singlets), and $\SUtwoL$ distinguishes the left-handed fermions (doublets) from the right-handed fermions (singlets). Finally, weak hypercharge further distinguishes SM particles according to the Gell-Mann--Nishijima formula, $Y=Q-I_3$, where $Q$ is electric charge and $I_3$ is the third component of weak isospin~\cite{sm_intro}.

    Experimentally, the $\Wpm$, $\Z$, quarks, and leptons are massive\fxnote{know about the weinberg angle, first measured precisely in neutrino electron elastic scattering}. Explicit mass terms for the fermions or gauge bosons would violate the chiral and gauge symmetries of the SM Lagrangian, however, so $\SUtwoLUoneY$ must be broken\fxnote{why can't you explicitly break gauge symmetry?}. Specifically, the nonzero vacuum expectation value (VEV)\fxnote{maybe just use $v$} of the Higgs field spontaneously breaks $\SUtwoLUoneY$ to $\UoneQ$. In the process, three of the four degrees of freedom associated with the complex Higgs doublet become the longitudinal polarizations of $\Wpm$ and $\Z$, giving them mass. The remaining degree of freedom is the physical Higgs boson. Expanding the Higgs field about its VEV produces Yukawa interactions such as those in Eq.~\ref{yukawa_terms}\fxnote{This still doesn't sound right to me. The Yukawa interactions aren't produced by anything other than somebody ad-hoc adding the terms to the Lagrangian. Masses for the gauge bosons are an inherent feature of the Higgs mechanism, masses for the fermions are an ad-hoc extension, says Andrew}. These terms are interpreted as giving mass to the quarks and charged leptons, where the mass of each particle is given by the product of the Higgs VEV, $v$, and the coupling constant, $y$~\cite{sm_intro}.

    \noindent \begin{equation}
        \mathcal{L} = -\frac{yv}{\sqrt{2}}\bar{\psi}\psi 
        \label{yukawa_terms}
    \end{equation}

    SM parameters, such as masses and couplings, are corrected by interactions with virtual particles through loop diagrams. Fig.~\ref{loop_diagrams}, for example, shows interactions that correct the Higgs and electron masses. Because the Higgs is a fundamental scalar, its corrections depend quadratically on the cutoff scale, $\Lambda$, which is the largest energy scale for which the SM is valid. The fermions and gauge bosons, on the other hand, only depend logarithmically\fxnote{be able to work out example} on $\Lambda$~\cite{dine_naturalness}. The next section explores the concept of naturalness, a major motivating principle for beyond the standard model (BSM) physics that is directly tied to the relationship between the bare Higgs mass and the terms that correct it.

@@

@@

    All massive particles correct the Higgs mass with terms proportional to $\Lambda^2$~\cite{dine_naturalness}\fxnote{show diagrams and terms?}. Experimentally, the Higgs mass is \SI{125}{\giga\electronvolt}~\cite{cms_higgs, atlas_higgs}, a value that must be equal to the sum of the bare mass and all corrections. Although corrections differ in sign, the precise cancellation of many extremely large terms to produce a value as small as the Higgs mass is extremely unlikely. In fact, the fine-tuning necessary to produce the Higgs mass is on the order of one part in \num{e30} if the SM is valid up to the Planck scale ($\mPlanck\approx$ \SI{e16}{\giga\electronvolt}), where quantum gravity becomes important~\cite{giudice_naturally}. While it is possible that the bare mass and all corrections just so happen to cancel so perfectly, such miraculous fine tuning could also be a hint that a deeper physical mechanism is at work\fxnote{be prepared to offer historical examples of times when fine tuning was a hint of some deeper theory and times when it was not}.

@@

    Supersymmetry (SUSY) introduces a new symmetry in which every SM field fits into a larger multiplet with an inherent symmetry between bosons and fermions\fxnote{I need to be better at explaining exactly what this means}. In its simplest form, SUSY results in one new boson for every SM fermion, one new fermion for every SM boson, and one new Higgs doublet.\fxnote{make sure I understand the new Higgs doublet} The increase in particle multiplicity calls for a new naming convention: the spin-0 superpartners to the SM fermions are referred to as sfermions (e.g. stop or smuon), and the spin-$\frac{1}{2}$ superpartners to the SM bosons are given the name of their SM counterpart with ``ino" tacked on to the end (e.g. Higgsino or wino).\fxnote{describe charginos, neutralinos?} Collectively the superpartners to the SM particles are known as sparticles~\cite{primer}.

@@

    \fxnote{Fix diagram} 
    \fxnote{Add particle labels} 
    \noindent \begin{figure}[htbp] \begin{center}
        \feynmandiagram [inline=(d.base), layered layout, medium, horizontal=b to c] {
            a [particle=\(H\)] -- [scalar] b,
            b -- [fermion, half left, edge label=\(t\)] c,
            c -- [fermion, half left] b,
            c -- [scalar] d [particle=\(H\)],
    };
    \quad + \quad
        \feynmandiagram [inline=(d.base), horizontal=a to b, layered layout, medium] {
            a [particle=\(H\)] -- [scalar] b [dot] 
            -- [out=135, in=45, loop, min distance=2.5cm, charged scalar, edge label=\(\tilde{t}\)] b 
            -- [scalar] c [particle=\(H\)],
        };
        \quad = \quad 0 \quad
        \caption{Corrections to the Higgs mass from the top (left) and stop (right) cancel in exact SUSY. Diagram created with~\cite{tikz}}
        \label{exact_susy_correction}
    \end{center} \end{figure}

@@

                \diagram* {
                    (n1a) -- [fermion] (n1b) -- [fermion] (p1),
                    (n2) -- [fermion] (p2),
                    (n3) -- [fermion] (p3),
                    (n1b) -- [boson, edge label=\(\W^{-}\)] (w),
                    (w) -- [fermion] (lc),
                    (ln) -- [fermion] (w),
                };
                \draw [decoration={brace}, decorate] (n3.south west) -- (n1a.north west)
                node [pos=0.5, left] {\(n\)};
                \draw [decoration={brace}, decorate] (p1.north east) -- (p3.south east)
                node [pos=0.5, right] {\(p\)};
            \end{feynman}
            \end{tikzpicture}
            \qquad\qquad
            \feynmandiagram [layered layout, horizontal=a to b] {
                a [particle=\(\mu\)] -- [fermion] b,
                b -- [fermion] d [particle=\(\nu_{\mu}\)],
                b -- [boson, edge label=\(\W^{-}\)] c,
                c -- [fermion] e [particle=\(e\)],
                c -- [anti fermion] f [particle=\(\nu_{e}\)],
            };
            \caption{Long-lived decays of the neutron and muon. Diagram created with~\cite{tikz}}
            \label{sm_llp_processes}
        \end{center} \end{figure}

    In SUSY models, LLPs arise for the same basic reasons, and different SUSY models produce LLPs in different ways. The most straightforward SUSY LLP occurs in models with conserved R-parity, a proposed symmetry that prevents proton decay~\cite{primer}\fxnote{Be prepared to show this. Maybe add some diagrams illustrating proton decay w/o R-parity}. In models with R-parity, the SM particles are assigned R-parity +1 and their superpartners are assigned R-parity -1. Conserving the product of R-parity at each vertex has two immediate phenomenological consequences: (1) sparticles can only be produced in pairs and (2) the lightest supersymmetric particle (LSP) is absolutely stable. A neutral LSP could show up in collider searches as missing energy ($\MET$), making $\MET$ a key signature in many conventional SUSY searches~\cite{primer}\fxnote{could mention DM as well}\fxnote{Andrew says define met and maybe explain how weakly interacting stable particles produce met}.

@@

    \fxnote{diagram or eqs for R-parity / RPV?}

@@

\section{The CMS Detector}\fxnote{Add pictures of everything}
    \subsection{Overview}
        \fxnote{define eta, coordinate system in general}
        \fxnote{Mention that ATLAS is similar, says Andrew}  
        \fxnote{Know about the beam, says stan}  
        The Compact Muon Solenoid (CMS) detector is designed to reconstruct a variety of particles from proton-proton collisions at the Large Hadron Collider (LHC)\fxnote{Make sure you are able to simply explain how the LHC works during your talk: how does it accelerate protons, how does it curve their trajectories in a circle, etc, says andrew}, a circular collider that ran with center-of-mass energy \SI{7}{\tera\electronvolt} in 2010-11, \SI{8}{\tera\electronvolt} in 2012, and \SI{13}{\tera\electronvolt} from 2015 through to the present\fxnote{cite something}. To reconstuct final-state particles, CMS uses a particle-flow algorithm that combines information from several subdetectors nested radially in and around a \SI{3.8}{\tesla} superconducting solenoid. All together, CMS is \SI{21.6}{\m} long, \SI{14.6}{\m} across, and weighs approximately \SI{12500}{t}~\cite{cms_experiment}. The subdetectors that make up CMS are shown in Fig.~\ref{cms} and described below.

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.7\textwidth]{images/cms}
        \caption{The CMS detector~\cite{cms_image}}
        \label{cms}
        \end{center} \end{figure}

@@

        \noindent \begin{figure}[htbp] \begin{center}
            \includegraphics[width=0.4\textwidth]{images/pixel_old_new}
            \caption{Diagram showing barrel pixel layers before and after the winter 2016/17 upgrade.~\cite{pixel_old_new}}
            \label{pixel_old_new}
        \end{center} \end{figure}

@@

    \subsection{Electromagnetic Calorimeter}
        \fxnote{I define EE \& EB but never use those definitions}
        After traversing the inner tracker, particles next encounter the electromagnetic calorimeter (ECAL). As a homogeneous scintillation calorimeter, ECAL uses \num{61200} lead tungstate crystals in the barrel and \num{7324} in each endcap to reconstruct the energy deposited during electromagnetic showers. Lead tungstate crystals allow for a fast (\SI{80}{\percent} of light within \SI{25}{\nano\s}), compact (radiation length = \SI{0.89}{cm}), fine-grained (Moli\`ere radius = \SI{2.2}{cm}), and radiation hard (up to 10 Mrad\fxnote{learn context for this number}) calorimeter. The main drawback is the relatively low light yield (\SI{30}{photon\per\mega\electronvolt}), which necessitates photodetectors with intrinsic gain that work in magnetic fields\cite{cms_experiment, cms_tdr}. Energy resolution is varies from approximately \num{1}-\SI{5}{\percent} depending on the amount of material traversed before reaching ECAL~\cite{ecal_performance}.

        The barrel section extends radially from \num{129} to \SI{177}{cm} and covers a pseudorapidity range of \num{0} to \num{1.479}. The crystals are tapered to approximately project back to the IP\fxnote{define if not already} but not so perfectly that likely particle trajectories align with cracks. Each crystal is approximately one Moli\`ere radius wide and 25.8 radiation lengths deep.\fxnote{Why do electrons and photons shower in the ECAL but not (usually) hadrons?, asks andrew}

        The crystals in each endcap section are arranged in an x-y grid that starts at $\lvert z \rvert = \SI{315}{cm}$ and covers a pseudorapidity range of \num{1.479} to \num{3.0}. Additionally, a preshower detector, which consists of two planes of silicon strip detectors and a lead absorber, sits just inside the grid of lead tungstate crystals. \fxnote{Say what the preshower is for? Is it for pi0 rejection? If so, how and why?}

    \subsection{Hadronic Calorimeter}
        \fxnote{define HB \& HF never to be used}
        The hadronic calorimeter (HCAL) uses \SI{3.7}{mm} plates of plastic scintillator interspersed within brass absorber to reconstruct the energy deposited during hadronic showers. Embedded wavelength-shifting fibers capture the scintillation light and transfer it to clear fibers to be read out by hybrid photodiodes\cite{cms_experiment, cms_tdr}.

        The barrel section (HB) spans a pseudorapidity range of \num{-1.4} to \num{1.4} and has \num{2304} towers \fxnote{check}, each of which contains \num{14} \fxnote{check} brass plates and covers an area of \num{0.087} x \SI{0.087}{in} $\eta-\phi$ \fxnote{get a feel for this area}. In addition, an extra layer (or two at $\eta = \num{0}$) of scintillator tiles sits just outside the solenoid. This extra layer, known as hadron outer (HO), spans a pseudorapidity range of \num{-1.26} to \num{1.26} and increases the minimum effective HCAL interaction length to greater than \num{11.8}.

        Each endcap spans a pseudorapidity range of \num{1.3} to \num{3.0} with \num{14} towers in eta and \num{5} to \SI{10}{\degree} $\phi$ segmentation. Also, a steel and quartz fiber forward calorimeter (HF) sits \SI{11.2}{m} from the interaction pipe in a pseudorapidity range of \num{3.0} to \num{5.0}. In HF, particles produce Cherenkov light when traversing the quartz fibers that run parallel to the beamline. \fxnote{say why HF is useful or mention crazy flux?}

@@

    \subsection{Muon System}
        \fxnote{Specifically mention CMS design specs to meet physics goals?}
        The muon system (MS), which is the outermost subdetector, detects muons with \num{3} varieties of gaseous ionization detectors: drift tubes (DTs) in the barrel, cathode strip chambers (CSCs) in the endcaps, and resistive plate chambers (RPCs) in both the barrel and endcaps. The choice of detector varies between barrel and endcap due to the difference in neutron-induced background\fxnote{what?} , muon rates, and magnetic field strength\fxnote{why does B matter?}. The RPCs are useful in the endcaps and barrel because their fast response and good time resolution allow unambiguous bunch crossing identification. Combining MS and tracker measurements improves the muon momentum resolution at high momenta where the tracker resolution declines (see Fig.~\ref{muon_pt_resolution}\cite{cms_experiment, cms_tdr}.

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.5\textwidth]{images/muon_pt_resolution}
            \caption{Transverse momentum resolution in the tracker and muon system for small (left) and large (right) pseudorapidity~\cite{cms_image}.}
        \label{muon_pt_resolution}
        \end{center} \end{figure}

@@

    \subsection{Trigger and Particle Reconstruction}
        \fxnote{Is entire detector read out at L1?}
        The trigger reduces the data writing rate from the \SI{40}{\mega\hertz} collision rate to less than \SI{1}{\kilo\hertz} so that events can be written to tape. The rate reduction happens in two stages: Level-1 (L1) and High-Level Trigger (HLT). L1 analyzes input from ECAL, HCAL, and MS with custom electronics to reduce the rate to $\approx$100 kHz in 3.8 us. With input from all subdetectors, the HLT then uses a dedicated processor farm to further reduce the rate to the desired $<$ 1 kHz\cite{cms_experiment, cms_trigger_upgrade}.

@@

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.7\textwidth]{images/jamie_cartoon}
        \caption{Diagram of several LLP topologies. Image courtesy of Jamie Antonelli}
        \label{jamie_cartoon}
        \end{center} \end{figure}

        Displaced vertex and disappearing track analyses both look for\fxnote{again, analysts look, not analyses} particles that are unstable on detector timescales. Displaced vertex searches are sensitive to LLPs that decay to visible particles within the inner volume \fxnote{specify} of the detector, leaving tracks that point back to a vertex some measurable distance from the nominal interaction point (IP)~\cite{atlas_displaced}. Disappearing track searches are sensitive to LLPs that decay to particles that are either neutral or too soft to be reconstructed~\cite{atlas_disappearing}. Together, these two types of analyses cover a wide range of lifetimes and topologies for LLPs that decay on detector timescales\fxnote{dangerous as hell. Be really good at giving a quantitative explanation of this sentence}\fxnote{find something to cite}.

@@

    \subsection{Displaced Vertices}
        LLPs that decay before reaching the inner tracker can be identified with a displaced vertex (DV)\fxnote{DVs defined clearly enough in overview?}\fxnote{Is it still a DV if it decays in active material somewhere?}. ATLAS performs a search for such particles by looking for at least one DV along with large MET and large track multiplicity~\cite{atlas_displaced}.\fxnote{make sure I can explain why met and multiplicity} The search is sensitive to LLPs with lifetimes in the \SI{20}{\pico\s} to \SI{10}{\nano\s} range\fxnote{Am I defining sensitivity range consistently across analyses?}.

@@

        \fxnote{Add particle labels} 
        \noindent \begin{figure}[htbp] \begin{center}
        \feynmandiagram [layered layout, horizontal=a to b] {
            a [particle=\(\tilde{g}\)] -- [plain, gluon] b,
            b -- [anti charged scalar, edge label=\(\tilde{q}\)] c,
            b -- [fermion] d [particle=\(q\)],
            c -- [anti fermion] e [particle=\(q\)],
            c -- [plain, boson] f [particle=\(\tilde{\chi}_1^0\)],
        };
        \caption{Long-lived gluino decay through highly virtual squark. Diagram created with~\cite{tikz}}
        \label{displaced_process}
        \end{center} \end{figure}

@@

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.6\textwidth]{images/material_map}
            \caption{Map of material dense regions of the ATLAS detector.\fxnote{explain more}~\cite{atlas_displaced}}
        \label{material_map}
        \end{center} \end{figure}

@@

    \subsection{Disappearing Tracks}
        Certain models with nearly mass-degenerate LSP and NLSP have the potential to present a remarkable signal: charged particles that seem to disappear in the tracker. Specifically, models with winos as the lightest gauginos\fxnote{have these words been defined?}, such as many AMSB models\fxnote{why amsb? what is amsb?}, will have a chargino NLSP with an observable lifetime (\SI{0.2}{\nano\s} is a typical estimate\fxnote{why? how typical?}) that eventually decays to the neutralino LSP and, commonly, a pion\fxnote{why?}. The nearly degenerate masses have two important phenomenological roles: (1) increase the chargino lifetime, and (2) ensure low-momentum decay products. The low-momentum pion is then unlikely to be reconstructed, which results in a disappearing track~\cite{atlas_disappearing}.

        Both ATLAS and CMS looked for disappearing tracks during Run I, but the tracker improvements installed in ATLAS before 2015 datataking and CMS before 2017 datataking should increase sensitivity to decays at shorter radii \fxnote{I haven't mentioned previous iterations when talking about other analyses}. The ATLAS search presented here is sensitive to LLP lifetimes of approximately \SI{10}{\pico\s} to \SI{10}{\nano\s} \fxnote{make sure I'm talking about lifetime sensitivity in the same way across analyses}.

>>>>

\newpage
\tableofcontents
\newpage
\doublespacing
%\linenumbers
\pagenumbering{arabic}

\section{The Standard Model} \label{SM}
    The standard model of particle physics (SM) describes all known particles and their non-gravitational interactions and is experimentally verified to astonishing accuracy. The SM consists of twelve spin-$\frac{1}{2}$ particles, the fermions, that make up all observed matter; twelve spin-1 particles, the gauge bosons, that communicate the electromagnetic, weak, and strong forces; and one fundamental scalar, the Higgs boson, which breaks electroweak symmetry, giving mass to the gauge bosons and fermions.

    The fermions and gauge bosons can be classified according to the forces with which they interact. The fermions are further divided into six quarks, which carry color and interact via the strong force, and six leptons, which do not. Furthermore, all six quarks and three leptons (the electron, muon, and tau) carry electric charge and interact electromagnetically. All fermions interact via the weak force. The gauge bosons are comprised of the photon, which communicates the electromagnetic force; the $\mathrm{W}^+$, $\mathrm{W}^-$, and $\mathrm{Z}$, which communicate the weak force; and eight gluons that communicate the strong force. Of these, only the $\mathrm{W}^+$ and $\mathrm{W}^+$ are electrically charged and only the gluons carry color charge.

@@

    SM parameters, such as masses and couplings, are corrected by interactions with virtual particles through loop diagrams. Fig.~\ref{loop_diagrams}, for example, shows interactions that correct the Higgs and electron masses. Because the Higgs is a fundamental scalar, its corrections depend quadratically on the cutoff scale, $\Lambda$, which is the largest energy scale for which the SM is valid. The fermions and gauge bosons, on the other hand, only depend logarithmically on $\Lambda$~\cite{dine_naturalness}. The next section explores the concept of naturalness, a major motivating principle for beyond the standard model (BSM) physics that is directly tied to the relationship between the bare Higgs mass and the terms that correct it.

@@

    All massive particles correct the Higgs mass with terms proportional to $\Lambda^2$~\cite{dine_naturalness}\fxnote{show diagrams and terms?}. Experimentally, the Higgs mass is \SI{125}{\giga\electronvolt}~\cite{cms_higgs, atlas_higgs}, a value that must be equal to the sum of the bare mass and all corrections. Although corrections differ in sign, the precise cancellation of many extremely large terms to produce a value as small as the Higgs mass is extremely unlikely. In fact, the fine-tuning necessary to produce the Higgs mass is on the order of one part in \num{e30} if the SM is valid up to the Planck scale ($\mPlanck\approx$ \SI{e16}{\giga\electronvolt}), where quantum gravity becomes important~\cite{giudice_naturally}. While it is possible that the bare mass and all corrections just so happen to cancel so perfectly, such miraculous fine tuning could also be a hint that a deeper physical mechanism is at work.

@@

    Supersymmetry (SUSY) introduces a new symmetry in which every SM field fits into a larger multiplet with an inherent symmetry between bosons and fermions. In its simplest form, SUSY results in one new boson for every SM fermion, one new fermion for every SM boson, and one new Higgs doublet. The increase in particle multiplicity calls for a new naming convention: the spin-0 superpartners to the SM fermions are referred to as sfermions (e.g. stop or smuon), and the spin-$\frac{1}{2}$ superpartners to the SM bosons are given the name of their SM counterpart with ``ino" tacked on to the end (e.g. Higgsino or wino).\fxnote{describe charginos, neutralinos?} Collectively the superpartners to the SM particles are known as sparticles~\cite{primer}.

@@

    \noindent \begin{figure}[htbp] \begin{center}
        \feynmandiagram [inline=(d.base), layered layout, medium, horizontal=b to c] {
            a [particle=\(H\)] -- [scalar] b,
            b -- [fermion, half left, edge label=\(t\)] c,
            c -- [fermion, half left] b,
            c -- [scalar] d [particle=\(H\)],
    };
    \quad + \quad
        \feynmandiagram [inline=(d.base), horizontal=a to b, layered layout, medium] {
            a [particle=\(H\)] -- [scalar] b [dot] 
            -- [out=135, in=45, loop, min distance=2.5cm, charged scalar, edge label=\(\tilde{t}\)] b 
            -- [scalar] c [particle=\(H\)],
        };
        \quad = \quad 0 \quad
        \caption{Corrections to the Higgs mass from the top (left) and stop (right) cancel in exact SUSY. Diagram created with~\cite{tikz}.}
        \label{exact_susy_correction}
    \end{center} \end{figure}

@@

                \diagram* {
                    (n1a) -- [fermion] (n1b) -- [fermion] (p1),
                    (n2) -- [fermion] (p2),
                    (n3) -- [fermion] (p3),
                    (n1b) -- [boson, edge label=\(\W^{-}\)] (w),
                    (w) -- [fermion] (lc),
                    (ln) -- [fermion] (w),
                };
                \draw [decoration={brace}, decorate] (n3.south west) -- (n1a.north west)
                node [pos=0.5, left] {\(n\)};
                \draw [decoration={brace}, decorate] (p1.north east) -- (p3.south east)
                node [pos=0.5, right] {\(p\)};
            \end{feynman}
            \end{tikzpicture}
            \qquad\qquad
            \feynmandiagram [layered layout, horizontal=a to b] {
                a [particle=\(\mu\)] -- [fermion] b,
                b -- [fermion] d [particle=\(\nu_{\mu}\)],
                b -- [boson, edge label=\(\W^{-}\)] c,
                c -- [fermion] e [particle=\(e\)],
                c -- [anti fermion] f [particle=\(\nu_{e}\)],
            };
            \caption{Long-lived decays of the neutron and muon. Diagram created with~\cite{tikz}.}
            \label{sm_llp_processes}
        \end{center} \end{figure}

    In SUSY models, LLPs arise for the same basic reasons, and different SUSY models produce LLPs in different ways. The most straightforward SUSY LLP occurs in models with conserved R-parity, a proposed symmetry that prevents proton decay~\cite{primer}. In models with R-parity, the SM particles are assigned R-parity +1 and their superpartners are assigned R-parity -1. Conserving the product of R-parity at each vertex has two immediate phenomenological consequences: (1) sparticles can only be produced in pairs and (2) the lightest supersymmetric particle (LSP) is absolutely stable. A neutral LSP could show up in collider searches as missing energy ($\MET$), making $\MET$ a key signature in many conventional SUSY searches~\cite{primer}\fxnote{could mention DM as well}\fxnote{Andrew says define met and maybe explain how weakly interacting stable particles produce met}.

@@

@@

\section{The CMS Detector}\fxnote{Add pictures of everything}
    \subsection{Overview}
        \fxnote{define eta, coordinate system in general}
        \fxnote{Mention that ATLAS is similar, says Andrew}  
        The Compact Muon Solenoid (CMS) detector is designed to reconstruct a variety of particles from proton-proton collisions at the Large Hadron Collider (LHC), a circular collider that ran with center-of-mass energy \SI{7}{\tera\electronvolt} in 2010-11, \SI{8}{\tera\electronvolt} in 2012, and \SI{13}{\tera\electronvolt} from 2015 through to the present\fxnote{cite something}. To reconstuct final-state particles, CMS uses a particle-flow algorithm that combines information from several subdetectors nested radially in and around a \SI{3.8}{\tesla} superconducting solenoid. All together, CMS is \SI{21.6}{\m} long, \SI{14.6}{\m} across, and weighs approximately \SI{12500}{t}~\cite{cms_experiment}. The subdetectors that make up CMS are shown in Fig.~\ref{cms} and described below.

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.7\textwidth]{images/cms}
        \caption{The CMS detector~\cite{cms_image}.}
        \label{cms}
        \end{center} \end{figure}

@@

        \noindent \begin{figure}[htbp] \begin{center}
            \includegraphics[width=0.4\textwidth]{images/pixel_old_new}
            \caption{Diagram showing barrel pixel layers before and after the winter 2016/17 upgrade~\cite{pixel_old_new}.}
            \label{pixel_old_new}
        \end{center} \end{figure}

@@

    \subsection{Electromagnetic Calorimeter}
        After traversing the inner tracker, particles next encounter the electromagnetic calorimeter (ECAL). As a homogeneous scintillation calorimeter, ECAL uses \num{61200} lead tungstate crystals in the barrel and \num{7324} in each endcap to reconstruct the energy deposited during electromagnetic showers. Lead tungstate crystals allow for a fast (\SI{80}{\percent} of light within \SI{25}{\nano\s}), compact (radiation length = \SI{0.89}{cm}), fine-grained (Moli\`ere radius = \SI{2.2}{cm}), and radiation hard (up to 10 Mrad) calorimeter. The main drawback is the relatively low light yield (\SI{30}{photon\per\mega\electronvolt}), which necessitates photodetectors with intrinsic gain that work in magnetic fields\cite{cms_experiment, cms_tdr}. Energy resolution is varies from approximately \num{1}-\SI{5}{\percent} depending on the amount of material traversed before reaching ECAL~\cite{ecal_performance}.

        The barrel section extends radially from \num{129} to \SI{177}{cm} and covers a pseudorapidity range of \num{0} to \num{1.479}. The crystals are tapered to approximately project back to the IP\fxnote{define if not already} but not so perfectly that likely particle trajectories align with cracks. Each crystal is approximately one Moli\`ere radius wide and 25.8 radiation lengths deep. The crystals in each endcap section are arranged in an x-y grid that starts at $\lvert z \rvert = \SI{315}{cm}$ and covers a pseudorapidity range of \num{1.479} to \num{3.0}.

    \subsection{Hadronic Calorimeter}
        The hadronic calorimeter (HCAL) uses \SI{3.7}{mm} plates of plastic scintillator interspersed within brass absorber to reconstruct the energy deposited during hadronic showers. Embedded wavelength-shifting fibers capture the scintillation light and transfer it to clear fibers to be read out by hybrid photodiodes\cite{cms_experiment, cms_tdr}.

        The barrel section (HB) spans a pseudorapidity range of \num{-1.4} to \num{1.4} and has \num{2304} towers \fxnote{check}, each of which contains \num{14} \fxnote{check} brass plates and covers an area of \num{0.087} x \SI{0.087}{in} $\eta-\phi$. In addition, an extra layer (or two at $\eta = \num{0}$) of scintillator tiles sits just outside the solenoid. This extra layer, known as hadron outer (HO), spans a pseudorapidity range of \num{-1.26} to \num{1.26} and increases the minimum effective HCAL interaction length to greater than \num{11.8}.

        Each endcap spans a pseudorapidity range of \num{1.3} to \num{3.0} with \num{14} towers in eta and \num{5} to \SI{10}{\degree} $\phi$ segmentation. Also, a steel and quartz fiber forward calorimeter (HF) sits \SI{11.2}{m} from the interaction pipe in a pseudorapidity range of \num{3.0} to \num{5.0}. In HF, particles produce Cherenkov light when traversing the quartz fibers that run parallel to the beamline.

@@

    \subsection{Muon System}
        The muon system (MS), which is the outermost subdetector, detects muons with \num{3} varieties of gaseous ionization detectors: drift tubes (DTs) in the barrel, cathode strip chambers (CSCs) in the endcaps, and resistive plate chambers (RPCs) in both the barrel and endcaps. The choice of detector varies between barrel and endcap due to the difference in neutron-induced background, muon rates, and magnetic field strength. The RPCs are useful in the endcaps and barrel because their fast response and good time resolution allow unambiguous bunch crossing identification. Combining MS and tracker measurements improves the muon momentum resolution at high momenta where the tracker resolution declines (see Fig.~\ref{muon_pt_resolution}\cite{cms_experiment, cms_tdr}.

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.6\textwidth]{images/muon_pt_resolution}
        \caption{Transverse momentum resolution in the tracker and muon system for small (left) and large (right) pseudorapidity~\cite{cms_image}.}
        \label{muon_pt_resolution}
        \end{center} \end{figure}

@@

    \subsection{Trigger}
        The trigger reduces the data writing rate from the \SI{40}{\mega\hertz} collision rate to less than \SI{1}{\kilo\hertz} so that events can be written to tape. The rate reduction happens in two stages: Level-1 (L1) and High-Level Trigger (HLT). L1 analyzes input from ECAL, HCAL, and MS with custom electronics to reduce the rate to $\approx$100 kHz in 3.8 us. With input from all subdetectors, the HLT then uses a dedicated processor farm to further reduce the rate to the desired $<$ 1 kHz\cite{cms_experiment, cms_trigger_upgrade}.

@@

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.7\textwidth]{images/jamie_cartoon}
        \caption{Diagram of several LLP topologies. Image courtesy of Jamie Antonelli.}
        \label{jamie_cartoon}
        \end{center} \end{figure}

        Displaced vertex and disappearing track analyses both look for\fxnote{again, analysts look, not analyses} particles that are unstable on detector timescales. Displaced vertex searches are sensitive to LLPs that decay to visible particles within the inner volume \fxnote{specify} of the detector, leaving tracks that point back to a vertex some measurable distance from the nominal interaction point (IP)~\cite{atlas_displaced}. Disappearing track searches are sensitive to LLPs that decay to particles that are either neutral or too soft to be reconstructed~\cite{atlas_disappearing}. Together, these two types of analyses cover a wide range of lifetimes and topologies for LLPs that decay on detector timescales\fxnote{find something to cite}.

@@

    \subsection{Displaced Vertices}
        LLPs that decay before reaching the inner tracker can be identified with a displaced vertex (DV)\fxnote{DVs defined clearly enough in overview?}. ATLAS performs a search for such particles by looking for at least one DV along with large MET and large track multiplicity~\cite{atlas_displaced}. The search is sensitive to LLPs with lifetimes in the \SI{20}{\pico\s} to \SI{10}{\nano\s} range\fxnote{Am I defining sensitivity range consistently across analyses?}.

@@

        \noindent \begin{figure}[htbp] \begin{center}
        \feynmandiagram [layered layout, horizontal=a to b, small] {
            a [particle=\(\tilde{g}\)] -- [plain, gluon] b,
            b -- [anti charged scalar, edge label=\(\tilde{q}\)] c,
            b -- [fermion] d [particle=\(q\)],
            c -- [anti fermion] e [particle=\(q\)],
            c -- [plain, boson] f [particle=\(\tilde{\chi}_1^0\)],
        };
        \caption{Long-lived gluino decay through highly virtual squark. Diagram created with~\cite{tikz}.}
        \label{displaced_process}
        \end{center} \end{figure}

@@

        \noindent \begin{figure}[htbp] \begin{center}
        \includegraphics[width=0.6\textwidth]{images/material_map}
        \caption{Map of material dense regions of the ATLAS detector\fxnote{explain more}~\cite{atlas_displaced}.}
        \label{material_map}
        \end{center} \end{figure}

@@

    \subsection{Disappearing Tracks}
        Certain models with nearly mass-degenerate LSP and NLSP have the potential to present a remarkable signal: charged particles that seem to disappear in the tracker. Specifically, models with winos as the lightest gauginos\fxnote{have these words been defined?}, such as many AMSB models, will have a chargino NLSP with an observable lifetime (\SI{0.2}{\nano\s} is a typical estimate that eventually decays to the neutralino LSP and, commonly, a pion\fxnote{why?}. The nearly degenerate masses have two important phenomenological roles: (1) increase the chargino lifetime, and (2) ensure low-momentum decay products. The low-momentum pion is then unlikely to be reconstructed, which results in a disappearing track~\cite{atlas_disappearing}.

        Both ATLAS and CMS looked for disappearing tracks during Run I, but the tracker improvements installed in ATLAS before 2015 datataking and CMS before 2017 datataking should increase sensitivity to decays at shorter radii. The ATLAS search presented here is sensitive to LLP lifetimes of approximately \SI{10}{\pico\s} to \SI{10}{\nano\s}.

        The authors consider two signal processes, both shown in Fig.\ref{disappearing_processes}: (1) electroweak gaugino production resulting in a disappearing track, an ISR jet, and MET and (2) gluino pair production resulting in a disappearing track, four jets, and MET. Process (1) could be the only gaugino production mode at the LHC if the gluino mass is kinematically inaccessible, but process (2), which leads to chargino production in cascade decays, will be dominant if the gluino mass is within reach.
